Stack Safety as a Security Property

### 2021-03-04 main.tex
Working on example for sharing
Small reorg
Updating annotation and context steps for sharing

\section{Examples and Intuition}
\input{example1}

@@

Separate from the machine state, our formal properties will keep track of additional context
information representing the history of the run. The following definitions are parameterized
over an arbitrary set of context states, \(\CONTEXTS\), an initial context state \(\context_0
\in \CONTEXTS\), and a context step function \(\mach, \context \constep \context' \in
\MACHS \times \CONTEXTS \rightarrow \CONTEXTS\).

@@

\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap\})\]
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \option{\ann}\]

@@

\definition A system enjoys {\em ultra-eager integrity} if for any reachable state
\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon{} \mpcstatename'\),
and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
\pi_\mach(\mpcstatename') ~ \component\).

@@

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold for
\(\depth' \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\depth),\_\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
Then take any \(\nach\) such that \(\mach \approx_\components \nach\). \(\MPCT\) corresponds to the
trace of a subroutine from its call to its return.

@@

\subsection{Stack Safety with Sharing}
\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap\} \cup
\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]

@@

@@

\section{Coroutine Safety}
\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap\} \cup
\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]

\subsection{Well-bracketed Control Flow}
\label{sec:wbcf}

>>>>

\section{Examples and Intuition}
\label{sec:running-example}

\input{example1}

\input{example2}

@@

Separate from the machine state, our formal properties will keep track of additional context
information representing the history of the run. The following definitions are parameterized
over an arbitrary set of context states, \(\CONTEXTS\), annotations \(\ANNS\),
an initial context state \(\context_0 \in \CONTEXTS\), and a context step function
\(\mach, \context, \ann \constep \context' \in \MACHS \times \CONTEXTS \times \ANNS \rightarrow \CONTEXTS\).

@@

\[\ann \in \ANNS ::= \callmap + \retmap\]
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \option(\ann)\]

@@

\definition A system enjoys {\em ultra-eager integrity} if for any reachable state
\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
\pi_\mach(\mpcstatename') ~ \component\).

@@

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold for
\(\depth' \geq \depth\). For any reachable segment \(\MPCT \langle f_\depth \rangle\),
let \(\head(\MPCT) = \mach,\pol,(\domMap,\depth)\) and
\(\components\) be the set of components \(\component\) such that \(\domMap ~ \component \not = \outside\).
Then take any \(\nach\) such that \(\mach \approx_\components \nach\). \(\MPCT\) corresponds to the
trace of a subroutine from its call to its return.

@@

@@

Having seen instances of eager confidentiality and integrity, defined in terms of reachable trace
segments, we can generalize them to properties of the trace segments themselves. Then as we expand
our model, our further properties will be defined in terms of the trace properties.

\definition Let \(\MPCT\) be a reachable segment and \(\components\) be any set of components.
\(\MPCT\) enjoys {\em trace integrity} when \(\last(\MPCT) ~ \component = \head(\MPCT) ~ \component\)
for all \(\component \in \components\), and when \(\last(\MPCT)\) is undefined.

\definition Let \(\MPCT \langle f \rangle\) be a reachable segment delimited by the predicate \(f\)
and the \(\components\) be any set of components. Let \(\head(\MPCT) = \mpcstate{\mach}{\pol}{\context}\)
and \(\nach\) be a machine state such that \(\mach \approx_\components \nach\). Suppose we take the
induced trace from \(\npcstatename = \mpcstate{\nach}{\pol}{\context}\) that corresponds to \(\MPCT\),
\(\npcstatename \hookrightarrow \NPCT | \neg f\). The trace \(\MPCT\) enjoys {\em trace confidentiality}
with respect to \(f\) and \(P\) if, for all such \(\NPCT\), they correspond according to the four cases
above:

\subsection{Stack Safety with Sharing}

Next we introduce sharing. We will need to extend the set of annotations to include
sharing and passing, and now the code map may map a single instruction to multiple
annotations, because in some calling conventions a single instruction might carry
meaning as both a call and a pass. But we will never have a single instruction annotated
as both a call and a return.

\[\ann \in \ANNS ::= \callmap, \retmap + (\sharemap ~ P \in \MACHS \times \COMPONENTS) +
(\passmap ~ P \in \MACHS \times \COMPONENTS)\]
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \powerset(\ann)\]

And we extend the set of domains to include shared and passed components, the latter of
which are labeled with the depth of the caller that has passed them.

\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \shared{\depth} + \sealed{\depth} + \passed{\depth}\]
\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
\times (\depth \in \mathbb{N}) \]

What it means to push the stack is similar to above, but two details are worth noting.
First, shared and passed components are not sealed by the sealing convention, and second, anything
not touched by the sealing convention must be unsealed. Sharing and passing cannot apply to memory
the callee could otherwise use.

\[
(\push{\domMap}{\depth}{\components}) ~ \component =
\begin{cases}
  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
                    \text{ and } \component \in \components \\
  \domMap ~ \component & \text{else if } \component \in \components \\
  \unsealed & \text{else} \\                  
\end{cases}
\]

Popping is simpler, but note that while passed components are unsealed on return to their
passer, shared ones are unsealed on return {\em from} their sharer - on deallocation, in other words.

\[
(\pop{\domMap}{\depth}) ~ \component =
\begin{cases}
  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth-1}
              \text{ or } \domMap ~ \component = \passed{\depth-1}
              \text{ or } \domMap ~ \component = \shared{\depth} \\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]

Now we can inductively define the context update function; as these could be ambiguous assume that
they apply in the order presented.

\begin{center}
  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \callmap \cup \passmap \cup \_\)}
              {\(\mach,(\domMap,\depth) \stepstocon (\push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1)\)}
\hspace*{0.1\textwidth}

  \judgmenttwo{\(\mach \stepsto \mach'\)}
              {\(\codemap ~ (\mach ~ \PCname) = \funid, \mathbf{Some} ~ \retmap\)}
              {\(\mach,(\domMap,\depth) \stepstocon \mach',(\pop{\domMap}{\depth},\depth-1)\)}
\end{center}

@@

\subsection{Coroutine Safety}
\[\ann \in \ANNS ::= \powerset(\{\callmap, \retmap, \yieldmap\} \cup
\{\sharemap ~ P | P \subseteq \MACHS \times \COMPONENTS \} \cup
\{\passmap ~ P | P \subseteq \MACHS \times \COMPONENTS \})\]

\section{Well-bracketed Control Flow}
\label{sec:wbcf}
